\documentclass[11pt]{extarticle}
\usepackage{fullpage,amsmath,amsfonts,microtype,nicefrac,amssymb, amsthm}
\usepackage[left=1in, bottom=1in, top=1in, right = 1in]{geometry}
\usepackage{textcomp}
\usepackage{mathpazo}
\usepackage{mathrsfs}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}

\usepackage{microtype}

\usepackage{bm}
\usepackage{dsfont}
\usepackage{enumerate}
\usepackage{ragged2e}

\setlength{\parindent}{24pt}
\setlength{\jot}{8pt}


\usepackage[shortlabels]{enumitem}


%% FOOTNOTES
\usepackage[bottom]{footmisc}
\usepackage{footnotebackref}


%% FIGURE ENVIRONMENT
%\graphicspath{{}}
\usepackage[margin=15pt, font=small, labelfont={bf}, labelsep=period]{caption}
\usepackage{subcaption}
\captionsetup[figure]{name={Figure}, position=above}
\usepackage{float}
\usepackage{epstopdf}


%% NEW COMMANDS
\renewcommand{\baselinestretch}{1.25} 
\renewcommand{\qedsymbol}{$\blacksquare$}
\newcommand{\R}{\mathbb{R}}
\newcommand{\indep}{\mathrel{\text{\scalebox{1.07}{$\perp\mkern-10mu\perp$}}}}
\renewcommand{\b}{\begin}
\newcommand{\e}{\end}

%% NEWTHEOREM
\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{ex}[thm]{Example}
\newtheorem{remark}[thm]{Remark}
\newtheorem{cor}[thm]{Corollary}

%% LINKS and COLORS
\usepackage[dvipsnames]{xcolor}
\usepackage{hyperref}
\definecolor{myred}{RGB}{163, 32, 45}
\hypersetup{
	%backref=true,
	%pagebackref=true,
	colorlinks=true,
	urlcolor=myred,
	citecolor=myred, 
	linktoc=all,     
	linkcolor=myred,
}

%% TABLE OF CONTENTS
\addto\captionsenglish{
	\renewcommand{\contentsname}
	{}% This removes the heading over the table of contents.
}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%            END PREAMBLE           %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Dynamic Optimization: Problem Set \#5}

\author{Andreas Schaab}

\date{Fall, 2022}



\begin{document}

\maketitle
\thispagestyle{empty}
\setcounter{page}{0}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{10mm}
\section*{Problem 1}

\textbf{Credit:} David Laibson (\url{https://projects.iq.harvard.edu/econ2010c/problem-sets-david-laibson})

Please solve Problem 1 (``a simple consumption problem'') of David's problem set (PSET) \#3.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{10mm}
\section*{Problem 2}

\textbf{Credit:} David Laibson (\url{https://projects.iq.harvard.edu/econ2010c/problem-sets-david-laibson})

Please solve Problem 2 (``true / false / uncertain'') of David's problem set (PSET) \#3.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{10mm}
\section*{Problem 3}

\textbf{Credit:} David Laibson (\url{https://projects.iq.harvard.edu/econ2010c/problem-sets-david-laibson})

Please solve Problem 3 (``three period hyperbolic discounting model'') of David's problem set (PSET) \#3.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{10mm}
\section*{Problem 4}

\textbf{Credit:} David Laibson (\url{https://projects.iq.harvard.edu/econ2010c/problem-sets-david-laibson})

Please solve Problem 4 (``a procrastination problem'') of David's problem set (PSET) \#3.





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{10mm}
\section*{Problem 5}

\textbf{Credit:} Pablo Kurlat

%\textbf{Gerard:} add Pablo's fiscal multiplier problem here


Consider the following model.
\begin{itemize}
\item The world lasts for two periods
\item There are two ty pes of households.
-$\gamma$ type $A$ households. Preferences:
$$
u\left(c_1, c_2\right)=\log \left(c_1\right)+\beta \log \left(c_2\right)
$$
$-1-\gamma$ type $B$ households. Consumption rule:
$$
c_1=a+b\left(Y_1+s T\right)
$$
- All households get the same income
\item $Y_1$ is period-1 GDP
\item $s T$ is a lump-sum transfer from the government in period $1 . s$ is a constant
\item Period 2 GDP $Y_2$ is exogenously given
- e.g. flexible prices and inelastic labor supply
\item Period 1 GDP $Y_1$ is demand-determined
\item The real interest $r$ is fixed at zero
- e.g. due to rigid prices and the zero lower bound
\item The government spends $G$ in period 1
\item The government charges a lump-sum tax $\tau=s T+G$ on everyone in period 2
\item Notation:
- Aggregate consumption:
$$
C_1 \equiv \gamma c_1^A+(1-\gamma) c_1^B
$$
- Direct effect of transfers on consumption
\end{itemize}



\begin{enumerate}[(a)]
\item Solve for consumption of A and B households and for total consumption. Use the following
notation

\begin{itemize}
\item $c_1^i\left(T, \tau, Y_1\right)$ for $i=A, B$ denotes the consumption choice given transfers, taxes and period-1 GDP.

\item $c_1^i\left(T, G, Y_1\right)$ for $i=A, B$ denotes the consumption choice given transfers, government spending and period-1 GDP, once we have taken into account the government budget constraint, which implies a certain level of $\tau$ for given levels of $T$ and $G$.


\item $C_1\left(T, \tau, Y_1\right) \equiv \gamma c_1^A\left(T, \tau, Y_1\right)+(1-\gamma) c_1^B\left(T, \tau, Y_1\right)$

\item $C_1\left(T, G, Y_1\right) \equiv \gamma c_1^A\left(T, G, Y_1\right)+(1-\gamma) c_1^B\left(T, G, Y_1\right)$

\end{itemize}

\item Compute the following quantities

(a) $\frac{d Y_1}{d G}$



(b) $\frac{\partial C_1\left(T, \tau, Y_1\right)}{\partial T}$


(c) $\frac{\partial C_1\left(T, G, Y_1\right)}{\partial T}$


(d) $\frac{\partial C_1\left(T, G, Y_1\right)}{\partial G}$


(e) $\frac{d Y_1}{d T}$


\item Consider the following empirical experiments:

(a) A period-1 transfer $T_i$ is assigned to household $i$ in period 1. $T_i$ varies randomly across the population. The government budget is balanced with a lump-sum tax paid equally by everybody in period 2 .

(b) A lump-sum transfer $T$ is assigned equally to everyone. A random sample of the population gets in in period 1 and the rest get it in period 2. The government budget is balanced with a lump-sum tax paid equally by everybody in period 2 .
In both cases a researcher estimates a regression:
$$
c_{i 1}=\alpha+\delta T_{i 1}
$$
What quantity is the researcher estimating in each case? What quantity are Parker et. al. (2013) estimating?



\item Consider a government-spending-multiplier design where we use exogenous variation in (nationallevel) military spending to measure the contemporaneous government spending multiplier. What quantity are we measuring?


\item  Consider the case where $\gamma=0$.

(e.1) Solve for the equilibrium level of $Y_1$ as a function of $T$ and $G$.

(e.2) Show that:
$$
\frac{d Y_1}{d T}=\frac{d Y_1}{d G} \frac{\partial C_1\left(T, \tau, Y_1\right)}{\partial T}
$$
and also:
$$
\frac{d Y_1}{d T}=\frac{d Y_1}{d G} \frac{\frac{\partial C_1\left(T, G, Y_1\right)}{\partial T}}{1+\frac{\partial C_1\left(T, G, Y_1\right)}{\partial G}}
$$

\item Now consider the case where $\gamma>0$

(a) Solve for the equilibrium level of $Y_1$ as a function of $T$ and $G$.



(b) Show that (2) does not hold but (3) does hold.

\item Suppose that we want to measure $\frac{d Y_1}{d T}$ by combining the estimates of Parker et. al. (2013) and estimates of the government spending multiplier. Will we get the right answer?


\end{enumerate}


\end{document}











